Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2021-08-23 article/draft.tex
typos and updates in the section Tw reducing the visibility

For climate change we found that most of the videos ($13$ out of $18$) did not display an information panel. Lastly, no information panels were found in the general health category set, that included videos that were not related to COVID-19 but contained misinformation concerning cures for cancer, abortion, and viruses. 
Therefore, it is likely that YouTube adds information panels below videos including controversial topics and that can contain misleading information like the climate change, COVID-19 and vaccines. 
It seems that YouTube is focusing its efforts on current ``hot'' topics with a potentially wide public impact (e.g. elections, COVID-19), leaving room for misinformation related to other topics to spread.

@@

Mainstream social media platforms can reduce the visibility of the content created or shared by specific users, whenever they violate the platforms' rules. The implementation of this policy varies across platforms and is not easy to verify ex-post. In what follows we provide means to verify this policy on Twitter and Facebook.

@@

One of Facebook’s measures to regulate misinformation is to reduce the spread of misleading content through their ranking system. Facebook ranks each post and/or ad by assigning to it a relevancy score, where a high score leads to a high likelihood of the post and/or the ad to appear on a user's newsfeed. Doing so, Facebook can make a post or a whole account less visible by decreasing the relevancy score of its content; this is precisely the $reduce$ measure (see the article on Facebook Newsroom (2018)~\cite{newsroom2}). This measure can be verified by looking at the number of views (reach) of a post, but this metric is not available via the APIs used to access Facebook data: CrowdTangle or Buzzsumo. Hence we can indirectly investigate the $reduce$ measure by looking at the engagement metrics (likes, comments, shares) related to a given post; which are available on CrowdTangle and Buzzsumo. If a post reaches less users because it has a lower ranking, then it is less likely to receive likes, comments and shares, relative to a post with a higher ranking.

@@

The number of public posts sharing an $Infowars$ link remained globally stable throughout $2019$ (see figure \ref{infowars1} top panel). Thus the measure announced by Facebook doesn't seem to have prevented users from sharing an $Infowars$ link. Nevertheless, a clear drop in engagement was observed on May 2, 2019 (see figure \ref{infowars1} bottom panel). The number of reactions, shares and comments per post have decreased respectively  by $-94\%$,  $-96\%$ and $-93\%$  when comparing the two months before and after May 2, 2019. This suggests the measure taken by Facebook in May $2019$, is not a ban per se, but rather a $reduce$ measure. This is because users could still post $Infowars$ links, but these posts generated less engagement. It should be noted that the engagement metrics increased again by the end of $2019$ / beginning of $2020$, suggesting that the $reduce$ measure may have been lifted a few months after its implementation.

Another measure that platforms can apply is to prevent users from sharing specific types of content, in this example URLs coming from a specific domain name. The Beauty of life (\href{https://thebl.com/}{thebl.com/}) is a US-based media company that shares pro-Trump views and conspiracy theories such as QAnon.
\footnote{ See \href{https://en.wikipedia.org/wiki/The\_Epoch\_Times\#Removal\_of\_The\_BL\_(The\_Beauty\_of\_Life)\_from\_Facebook}{wikipedia article}.} 
Facebook has announced on December $20$, $2019$ that ``The BL is now banned from Facebook''.
\footnote{\url{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us/}}

To verify Facebook’s ban, we first tested whether we could post a Facebook message containing a url from thebl.com. This turned out to be impossible. But such manual verification cannot inform us whether the ban applies indeed to all Facebook users and accounts (as we used only our own personal accounts), nor when it has started. To further investigate this policy, we collected data from the Buzzsumo API\footnote{BuzzSumo is a commercial content database that tracks the volume of user interactions with internet content on Facebook, Twitter, and other social media platforms.}. We used the ``/search/articles'' endpoint to collect the engagement metrics of the $13$ $634$ articles crawled from the $thebl.com$ website between January $1$, $2019$ and June $15$, $2021$.

@@

Using Buzzsumo data, we ascertained that links from thebl.com were not shared on Facebook anymore. The ban started on December $1$, $2019$, and appeared to be still enforced in June 2021.

@@

Twitter can take action against a tweet which violates the Twitter rules, by limiting its visibility\footnote{See the paragraph {\it Limiting Tweet visibility}: \href{https://help.twitter.com/en/rules-and-policies/enforcement-options}{https://help.twitter.com/en/rules-and-policies/enforcement-options}.} on users' timelines and in search results. To illustrate we provide an example for the website $globalresearch.ca$, which has several failed fact-checks according to \href{https://open.feedback.org/media/AEKA}{Feedback.org} and \href{https://iffy.news}{iffy.news} - two websites which provides databases of websites with low factual reporting levels.\footnote{For $globalresearch.ca$ see \href{https://open.feedback.org/media/AEKA}{open.feedback.org/media/AEKA} and \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/}.}

@@

%To illustrate we provide two examples linked to the Twitter accounts of the websites $globalresearch.ca$ and $off$-$guardian.org$. Both websites have several failed fact-checks according to \href{https://iffy.news}{iffy.news} - a website which provides a database of websites with low factual reporting levels.\footnote{For $globalresearch.ca$ see \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/} and for $off-guardian.org$ see \href{https://mediabiasfactcheck.com/offguardian/}{https://mediabiasfactcheck.com/offguardian/}.  }
\begin{figure}[h]
	\centering
		%\begin{multicols}{1}	
					\includegraphics[scale=0.35]{./img/globalresearch_14_06_2021_16pm_UTC.png} 
		%\end{multicols}
		%\begin{multicols}{1}
					\includegraphics[scale=0.32]{./img/globalresearch/tweet.png} 
		%\end{multicols}
		\caption{Screenshots taken on June 14, 2021. Left Panel: screenshot that shows that no results can be found when searching for $globalresearch.ca$ via the Twitter search box. Right Panel: example of a tweet posted on June $14$ containing the domain name {\it globalresearch.ca}. }
		\label{fig3}
\end{figure}

@@

The website $globalresearch.ca$ is linked to the Twitter account {$@CRG\_CRM$}. When a user searches via the twitter search-box for this domain name, no results appear as shown in the screenshot in the left panel of figure \ref{fig3}, taken on June $14$, $2021$. To further investigate the possible implementation of a reduced visibility measure, we search via the Twitter API for tweets, excluding retweets, containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $10$, $2021$.  As shown in the left panel in figure \ref{fig4}, we find a strictly positive number of tweets containing the domain name {\it globalresearch.ca} throughout May $2021$ and the first week of June $2021$. Hence, the visibility of tweets containing this domain name has been reduced because users can no longer access tweets containing the domain name {\it globalresearch.ca} via the search box. Nevertheless users are not restrained from posting tweets containing this domain name, as shown in the screenshot in panel $(b)$ of figure \ref{fig3}, found by taking the tweet ID of one of the collected tweets via the Twitter API. Furthermore, those collected Tweets have strictly positive engagement metrics as shown in panel $(b)$ of figure \ref{fig4}. Hence, the users who tweet articles from the {\it globalresearch.ca} website receive tweet level engagement from their own followers.

\begin{figure}[h]
\centering
		\includegraphics[scale=0.27]{../figure/globalresearch_domain_2021-08-17.jpg} 
		\includegraphics[scale=0.27]{../figure/globalresearch_domain_engagement_2021-08-17.jpg}
\caption{Left Panel: daily number of Tweets, excluding retweets, containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $30$, $2021$. Right Panel: engagement metrics of Tweets containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $30$, $2021$. Data collected via the Twitter API v2 on July $30$, 2021.   }
\label{fig4}
\end{figure}

Finally, the above mentioned account was recently suspended from Twitter. We noticed the message about the account suspension (see the right panel in figure \ref{fig4bis}) on May $25$, 2021. To the best of our knowledge, no official communication by Twitter has announced the suspension nor the exact date at which it was implemented. Hence the account may have gotten suspended before that date. Furthermore when a user attempts to click on a link which contains {\it globalresearch.ca} in a Tweet, a warning message appears and indicates that the link may be unsafe (see the left panel in figure \ref{fig4bis}). For the case of this Twitter account, it is likely that a mix of interventions was used: reducing the visibility via the search box, suspending the related Twitter account and returning a warning message when users click on a link containing this domain name.

>>>>

For climate change we found that most of the videos ($13$ out of $18$) did not display an information panel. Lastly, no information panels were found in the general health category set, that included videos that were not related to COVID-19 but contained misinformation concerning cures for cancer, abortion, and viruses. 
Therefore, it is likely that YouTube adds information panels below videos including controversial topics and that can contain misleading information like climate change, COVID-19 and vaccines. 
It seems that YouTube is focusing its efforts on current ``hot'' topics with a potentially wide public impact (e.g. elections, COVID-19), leaving room for misinformation related to other topics to spread.

@@

Mainstream social media platforms can reduce the visibility of content created or shared by users, whenever they violate the platforms' rules. 
The implementation of this policy varies across platforms and is not easy to verify ex-post. In what follows we provide indirect methods to explore how reducing the visibility of content works on Facebook, Twitter and YouTube.

@@

To tackle misinformation, Facebook can reduce the spread of misleading content through their built-in ranking system. 
More specifically, Facebook ranks each post and/or ad by assigning to it a relevancy score, where a high score leads to a high likelihood of the post and/or the ad to appear on a user's newsfeed. Doing so, Facebook can make a post or a whole account less visible by decreasing the relevancy score of its content; this is precisely the $reduce$ measure (see the article on Facebook Newsroom (2018)~\cite{newsroom2}). This measure can be verified by looking at the number of views (reach) of a post, but this metric is not available via the APIs used to access Facebook data: CrowdTangle or BuzzSumo. Hence we can indirectly investigate the $reduce$ measure by looking at the engagement metrics (likes, comments, shares) related to a given post; which are available on CrowdTangle and BuzzSumo. If a post reaches less users because it has a lower ranking, then it is less likely to receive likes, comments and shares, relative to a post with a higher ranking.

@@

The number of public posts sharing an $Infowars$ link remained globally stable throughout $2019$ (see figure \ref{infowars1} left panel). 
Thus the measure announced by Facebook does not seem to have prevented users from sharing an $Infowars$ link. Nevertheless, a clear drop in engagement was observed on May $2$, $2019$ (see figure \ref{infowars1} right panel). 
The number of reactions, shares and comments per post have decreased respectively  by $-94\%$,  $-96\%$ and $-93\%$  when comparing the two months before and after May $2$, $2019$. This suggests the measure taken by Facebook in May $2019$, is not a ban per se, but rather a $reduce$ measure. This is because users could still post $Infowars$ links, but these posts generated less engagement. It should be noted that the engagement metrics increased again by the end of $2019$ / beginning of $2020$, suggesting that the $reduce$ measure may have been lifted a few months after its implementation.

Another measure that platforms can apply is to prevent users from sharing specific types of content, like URLs belonging to a specific domain name.  
The Beauty of life (\href{https://thebl.com/}{thebl.com/}) is a US-based media company that shares pro-Trump views and conspiracy theories such as QAnon.
\footnote{ See \href{https://en.wikipedia.org/wiki/The\_Epoch\_Times\#Removal\_of\_The\_BL\_(The\_Beauty\_of\_Life)\_from\_Facebook}{wikipedia article}.} 
Facebook has announced on December $20$, $2019$ that ``The BL is now banned from Facebook''.
\footnote{\url{https://about.fb.com/news/2019/12/removing-coordinated-inauthentic-behavior-from-georgia-vietnam-and-the-us/}}

To verify Facebook’s ban, we first tested whether we could post a Facebook message containing a url from thebl.com. This turned out to be impossible. But such manual verification cannot inform us whether the ban applies indeed to all Facebook users and accounts (as we used only our own personal accounts), nor when it has started. To further investigate this policy, we collected data from the BuzzSumo API\footnote{BuzzSumo is a commercial content database that tracks the volume of user interactions with internet content on Facebook, Twitter, and other social media platforms.}. We used the ``/search/articles'' endpoint to collect the engagement metrics of the $13$ $634$ articles crawled from the $thebl.com$ website between January $1$, $2019$ and June $15$, $2021$.

@@

Using BuzzSumo data, we ascertained that links from thebl.com were not shared on Facebook anymore. The ban started on December $1$, $2019$, and appeared to be still enforced in June 2021.

@@

Twitter can take action against a tweet which violates the Twitter rules, by limiting its visibility\footnote{See the paragraph {\it Limiting Tweet visibility}: \href{https://help.twitter.com/en/rules-and-policies/enforcement-options}{https://help.twitter.com/en/rules-and-policies/enforcement-options}.} on users' timelines and in search results. To illustrate we provide an example for the website $globalresearch.ca$, which has several failed fact-checks according to \href{https://open.feedback.org/media/AEKA}{Feedback.org} and \href{https://iffy.news}{iffy.news} - two websites which provide databases of websites with low factual reporting levels.\footnote{For $globalresearch.ca$ see \href{https://open.feedback.org/media/AEKA}{open.feedback.org/media/AEKA} and \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/}.}

@@

%To illustrate we provide two examples linked to the Twitter accounts of the websites $globalresearch.ca$ and $off$-$guardian.org$. Both websites have several failed fact-checks according to \href{https://iffy.news}{iffy.news} - a website which provides a database of websites with low factual reporting levels.\footnote{For $globalresearch.ca$ see \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/} and for $off-guardian.org$ see \href{https://mediabiasfactcheck.com/offguardian/}{https://mediabiasfactcheck.com/offguardian/}.  }
\begin{figure}[h]
	\centering
		%\begin{multicols}{1}	
					\includegraphics[scale=0.35]{./img/globalresearch_14_06_2021_16pm_UTC.png} 
		%\end{multicols}
		%\begin{multicols}{1}
					\includegraphics[scale=0.32]{./img/globalresearch/tweet.png} 
		%\end{multicols}
		\caption{Screenshots taken on June 14, 2021. Left Panel: screenshot that shows that no results can be found when searching for articles linked to the domain name $globalresearch.ca$ via the Twitter search box. Right Panel: example of a tweet posted on June $14$ containing the domain name {\it globalresearch.ca}. }
		\label{fig3}
\end{figure}

@@

When a user searches via the twitter search-box for this domain name, no results appear as shown in the screenshot in the left panel of figure \ref{fig3}, taken on June $14$, $2021$. To further investigate the possible implementation of a reduced visibility measure, we searched via the Twitter API v2 for tweets, excluding retweets, containing the query {\it globalresearch.ca} from April $15$, $2021$ until August $15$, $2021$.  As shown in the left panel in figure \ref{fig4}, we find a strictly positive number of tweets containing the domain name {\it globalresearch.ca} after the date on which the screenshot was taken (left panel of figure \ref{fig3}). Hence, the visibility of tweets containing this domain name has been reduced because users can no longer access tweets containing the domain name {\it globalresearch.ca} via the search box. Nevertheless users are not restrained from posting tweets containing this domain name, as shown in the screenshot in panel $(b)$ of figure \ref{fig3}, found by taking the tweet ID of one of the collected tweets via the Twitter API. Furthermore, those collected Tweets have strictly positive engagement metrics as shown in panel $(b)$ of figure \ref{fig4}. Hence, the users who post tweets including articles from the {\it globalresearch.ca} website receive tweet level engagement from their own followers.

\begin{figure}[h]
\centering
		\includegraphics[scale=0.27]{../figure/globalresearch_domain_2021-08-17.jpg} 
		\includegraphics[scale=0.27]{../figure/globalresearch_domain_engagement_2021-08-17.jpg}
\caption{Left Panel: daily number of Tweets, excluding retweets, containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $30$, $2021$. Right Panel: engagement metrics of Tweets containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $30$, $2021$. Data collected via the Twitter API v2 on August $17$, 2021.   }
\label{fig4}
\end{figure}

Finally, the website $globalresearch.ca$ is linked to the Twitter account {$@CRG\_CRM$} and it was recently suspended from Twitter. We noticed the message about the account suspension (see the right panel in figure \ref{fig4bis}) on May $25$, 2021. To the best of our knowledge, no official communication by Twitter has announced the suspension nor the exact date at which it was implemented. Hence the account may have gotten suspended before that date. Furthermore when a user attempts to click on a link which contains {\it globalresearch.ca} in a Tweet, a warning message appears and indicates that the link may be unsafe (see the left panel in figure \ref{fig4bis}). For the case of this Twitter account, it is likely that a mix of interventions was used: reducing the visibility via the search box, suspending the related Twitter account and returning a warning message when users click on a link containing this domain name.
